---
title: "Colombia's Peace Agreement voting"
author: "Patrick Machado"
date: "5 jul 2019"
output: ioslides_presentation
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## BBC News, 3 oct 2016


*Voters in Colombia have rejected a landmark peace deal with Farc rebels in a shock referendum result, with 50.2% voting against it.*

*The deal was signed last week by President Juan Manuel Santos and Farc leader Timoleon Jimenez after nearly four years of negotiations.*

*But it needed to be ratified by Colombians in order to come into force.*

*Addressing the nation, President Santos said he accepted the result but would continue working to achieve peace.*

*[BBC News, 3 oct 2016](https://www.bbc.com/news/world-latin-america-37537252)*

---

```{r p_votes, echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}

# A. Datos
# Leer datos
Colombia.vote <- read.csv('Colombia.vote.csv')


# B. Back to back chart


# load packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(readr)

# prepare data for plotting
plotting_df <- Colombia.vote %>%  
    # select(departamento, si, no) %>% 
    transmute(Departamento = departamento,
           Si = parse_number(as.character(si), locale = 
                                 locale(grouping_mark = ".")), 
           No = -1 *parse_number(as.character(no), locale = 
                                locale(grouping_mark = "."))
           ) %>% 
    gather(key = "Vote", value = "Count", -Departamento)

plotting_df2 <- Colombia.vote %>%  
    # select(departamento, si, no) %>% 
    transmute(Departamento = departamento,
              Si = si.porc, 
              No = -1 * no.porc ) %>% 
    gather(key = "Vote", value = "Percentaje", -Departamento)

plotting_df$Percentaje <- plotting_df2$Percentaje


## find the order
temp_df <-
    plotting_df[plotting_df != 'COLOMBIA', ] %>% 
    filter(Vote == "Si") %>% 
    arrange(Count)
the_order <- temp_df$Departamento

# plot
p <- 
    plotting_df[plotting_df != 'COLOMBIA', ] %>% 
    ggplot(aes(x = Departamento, y = Count, group = Vote, 
               fill = Vote)) +
    geom_bar(stat = "identity", width = 0.75) +
    coord_flip() +
    scale_x_discrete(limits = the_order) +
    # another trick!
    scale_y_continuous(breaks = seq(-1500000, 1500000, 500000), 
                       labels = abs(seq(-1500000, 1500000, 500000))) +
    labs(x = "", y = "Votes", title = "Votes for each State") +
    theme(legend.position = "bottom",
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          panel.background = element_rect(fill =  "grey90")) #+
    # reverse the order of items in legend
    # guides(fill = guide_legend(reverse = TRUE)) +
    # change the default colors of bars
    # scale_fill_manual( values=c("red", "blue"),
    #                   name="",
    #                   breaks=c("No", "Si"),
    #                   labels=c("No", "Si")) 

# print(p)

## find the order for percentaje
temp_df <-
    plotting_df[plotting_df != 'COLOMBIA', ] %>% 
    filter(Vote == "Si") %>% 
    arrange(Percentaje)
the_order <- factor(c(as.character(temp_df$Departamento), 
                      'COLOMBIA'))

# percentaje plot
pp <- 
    plotting_df %>% 
    ggplot(aes(x = Departamento, y = Percentaje, group = Vote, 
               fill = Vote)) +
    geom_bar(stat = "identity", width = 0.75) +
    coord_flip() +
    scale_x_discrete(limits = the_order) +
    # another trick!
    scale_y_continuous(breaks = seq(-100, 100, 25), 
                       labels = abs(seq(-100, 100, 25))) +
    labs(x = "", y = "Votes %", title = 
             "Votes % distribution for each State") +
    theme(legend.position = "bottom",
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          panel.background = element_rect(fill =  "grey90")) #+
# reverse the order of items in legend
# guides(fill = guide_legend(reverse = TRUE)) +
# change the default colors of bars
# scale_fill_manual( values=c("red", "blue"),
#                   name="",
#                   breaks=c("No", "Si"),
#                   labels=c("No", "Si")) 

# print(pp)

# C. Plotly
ggplotly(p)


```

---


```{r p_percentaje, echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}

# C. Plotly
ggplotly(pp)

```

---

#### Votes for each State

```{r p_map, echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}

library(leaflet)

# Colombia.vote <- read.csv('Colombia.vote.csv')

# C. Iconos 

icono.si <- makeIcon(
    iconUrl = "secure-shield.png",
    iconWidth = 31*215/230, iconHeight = 31,
    iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

icono.no <- makeIcon(
    iconUrl = "banned.png",
    iconWidth = 31*215/230, iconHeight = 31,
    iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

icono.Colombia <- makeIcon(
    iconUrl = "banned.png",
    iconWidth = 42*215/230, iconHeight = 42,
    iconAnchorX = 42*215/230/2, iconAnchorY = 21
)

ajustes <- Colombia.vote$winner == 'Yes'
ajustes2 <- Colombia.vote$departamento == 'COLOMBIA'

# Map
peace <- leaflet() %>% 
    addTiles() %>%
    addMarkers(data = Colombia.vote[ajustes, ],
               popup = Colombia.vote$legend[ajustes], 
               icon = icono.si)  %>%
    addMarkers(data = Colombia.vote[!ajustes & !ajustes2, ],
               popup = Colombia.vote$legend[!ajustes & !ajustes2], 
               icon = icono.no) %>%
    addMarkers(data = Colombia.vote[ajustes2, ],
               popup = Colombia.vote$legend[ajustes2], 
               icon = icono.Colombia) 

peace

```

## Thanks!

Sources:

- Voting results: [data.world/juanbeta](https://data.world/juanbeta/colombia-votacin-por-la-paz-si-no-por-departamento)

- Colombia's geographical information: [dane.gov.co](https://geoportal.dane.gov.co/consultadivipola.html)

- Icons made by [Freepik](https://www.flaticon.com/authors/freepik) from [www.flaticon.com](https://www.flaticon.com/) is licensed by [CC 3.0 BY](http://creativecommons.org/licenses/by/3.0/)

- News article: [BBC News, 3 oct 2016](https://www.bbc.com/news/world-latin-america-37537252)


